---
title: "Data manipulation"
author: "Yan Li"
date: "2024-10-01"
output: github_document
---

```{r}
library(tidyverse)
```
# tidyverse is a collection of packages

# load FAS litters data and clean names
```{r}

litters_df = 
    read_csv("data/FAS_litters.csv", na = c("NA", ".", ""))
litters_df = 
    janitor::clean_names(litters_df)
```

## Select
choose some columns and not others
```{r}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)
```
### specify a range
```{r}
select(litters_df, group:gd_of_birth)
```

### remove variable
```{r}
select(litters_df, -pups_survive)
```

### rename variables
```{r}
select(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```
### rename all
```{r}
rename(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

### read about all of them, use start_with(),ends_with, contains()
```{r}
select(litters_df, starts_with("gd"))
```

### reorganizing columns without discarding anything
```{r}
select(litters_df, litter_number, pups_survive, everything())
```
### similar with relocate: to the beginning
```{r}
relocate(litters_df, litter_number, pups_survive)
```
# Filter
```{r}
filter(litters_df,gd0_weight < 22)
```

```{r}
filter(litters_df,gd0_weight == 20)
```

## not equal to 20
```{r}
filter(litters_df,!gd0_weight == 20)
```

## run both
```{r}
filter(litters_df,gd0_weight >= 22,gd_of_birth == 20 )
```
## either option

```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

# Mutate

## creates a new variable  measuring the difference between gd18_weight and gd0_weight and modifies the existing group variable(make lower case).
```{r}
mutate(litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```
# arrange(not used often)

```{r}
head(arrange(litters_df, group, pups_born_alive), 10)
```

# /> then

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  select(-pups_survive) |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) |> 
  drop_na(wt_gain)
```

# '%>%' %>% 
```{r}
litters_data_raw = read_csv("data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_selected, wt_gain = gd18_weight - gd0_weight)
```








