---
title: "tidy data"
author: "Yan Li"
date: "2024-10-01"
output: github_document
---
```{r}
library(tidyverse)
```

# pivot_longer
```{r}
pulse_df = 
  haven::read_sas("data/public_pulse_data.sas7bdat") |>
  janitor::clean_names()
pulse_df
```
-  the BDI score is spread across four columns, which correspond to four observation times. We can fix this problem using pivot_longer:
```{r}
pulse_tidy_df = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    values_to = "bdi")

pulse_tidy_df
```
- now visit is an issue. The original column names were informative but we probably don’t need to keep the bdi_score_ prefix in each case. I’ll use an additional option in pivot_longer to address this:
```{r}
pulse_tidy_df = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi")

pulse_tidy_df
```

- changing bl to 00m for consistency across visits and converting visit to a factor variable. (It’s possible that you would want visit to be a numeric variable instead, which could be done with a different call to mutate.)
```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)) 

print(pulse_df, n = 12)
```



